### 2018-06-29 Manuscript/carbonStabilizationMangrovesV2.tex
Worked on manuscript text

Here, our main objective was to determine the origin of the relatively high carbon levels in these soils using elemental and isotopic analyses of carbon as well as analyses of the soils' mineralogy in concert with Al- and Fe-oxide measurements. Analyses of stable isotopes are particularly useful to identify the origin of carbon in soils of coastal areas \citep{Bouillion, Spohn2012, Spohn2013}. In particular, we expect that: 1) the isotopic composition ($\updelta ^{13}$C) of soil carbon provides information on whether riverine sediments are a main source of C in Cisipat\'a bay, Colombia; and  2) the mineralogical composition of the sediments provides additional information on the origin of the soil carbon and the potential that it is mostly stabilized on the surface of Fe- and Al-oxides. Our main hypothesis is that soils in the basin mangroves are composed mainly by sediments transported by the Sin\'u river and deposited in the delta region, therefore explaining the relatively large values of TOC stored in these soils.

@@

\subsection{Field sampling}
Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}. We sampled 10 plots of 500 m$^2$ up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter (Eijkelkamp bi-partite gouge auger 04.03, Giesbeek, The Netherlands). We selected five randomly chosen points within each plot to extract soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4). In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig. \ref{fig:1}). Generally, cores without layer changes were divided into sections every 20 cm to explore differences of carbon concentration with depth. We divided conspicuous layer changes at the boundary, except for the sand and river sediment samples that were analyzed as one single sample for mineralogical and elemental composition.

@@

%\paragraph{X-ray diffraction analysis}
We conducted X-ray diffraction (XRD) measurements for qualitative and quantitative phase analyses \citep{spiess2009} on 12 representative samples including 4 end-member samples, and 4 samples of basin and fringe mangroves each. Samples were measured 20 min each, from 5 to 60 $^{\circ}2\uptheta$. We determined each mineral phase using the powder diffraction file (PDF) data. We also conducted a Rietveld refinement using the Topas software for XRD analysis (Bruker Corporation). To define the type of clays included in the samples, we further measured each sample before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels. XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.

@@

\subsection{Data analysis}
We performed statistical tests to determine differences in organic carbon concentrations, total nitrogen, $\delta^{13}$C, Fe- and Al-oxides, among the fringe and basin mangroves. After an initial test for normality (Shapiro-Wilk test), we found little evidence to support the normality assumption required in common tools such as ANOVA or t-tests. We therefore, performed the non-parametric Mann-Whitney U test, which is equivalent to the two-sample Wilcoxon test \citep{Hollander2015} under the null hypothesis that the samples from the fringe and basin mangroves  differ by a shift location $\mu =0$, and the alternative hypothesis that they different by some other shift location. In other words, the test helps to determine whether the samples from the two different mangrove types belong to the same statistical distribution as stated in the null hypothesis. Data and code to reproduce all results presented here can be obtained from the following repository \url{https://github.com/crlsierra/mangroveCstabilization.git}.

@@

Differences in bulk densities between the two forest types, multiplied by \% organic carbon, resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}). A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}). Summed across the profile, fringe mangroves had a TOC of 387.77 $\pm$ 125.77 Mg C ha$^{-1}$, and basing mangroves an average TOC of 665.41 $\pm$ 434.65 Mg C ha$^{-1}$. Given this large variability, we did not find enough evidence to reject the hypothesis that TOC from these two forest types are sampled from the same distribution  ($p$-value $= 0.47$, $W = 350$). TOC decreased with soil depth within the first four (basin) and three (fringe) layers. There was an increase in TOC for the last measured depth intervals of basin and fringe soils, which can be traced back to higher \%OC values of those layers.

We measured a higher proportion of negative $\updelta^{13}$C values for fringe soils \mbox{(-28 to -29 \permil)} than for basin mangroves soils \mbox{(-25 to -29 \permil)} (Fig. \ref{fig:3}).  Depth intervals 20-40 and 40-60 cm of basin soils show outliers with equal $^{13}$C values compared to the Sin\'{u} river end-member samples with values of around \mbox{-25 \permil}. The sand end-member samples have the most positive $^{13}$C values with approximately -3 \permil.  Strong variations in $^{13}$C are conspicuous for the depth intervals 20-40 and 40-60 cm of basin mangrove boxplots. We found no identifiable continuous trend between $^{13}$C values and depth. However, differences in $^{13}$C mean values between fringe and basin mangroves were statistically significant ($p$-value $< 0.001$, $F = 24.35$, 49 d.f.).

There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}). The proportion of mineral soil is larger in basin than in fringe mangroves. Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.

@@

The oxalate extraction dissolved much of the poorly crystalline Fe and Al oxides from the amorphous materials, whereas the dithionite extraction dissolved the crystalline Fe oxides as well as the amorphous materials \mbox{(Fig. \ref{fig:6})}. Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $< 0.001$, $F = 16.35$, 39 d.f.). Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $F = 33.45$, 39 d.f.).
Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values. While riverine samples had similar Al and Fe concentrations than those of basin mangroves in all 4 extraction patterns, metal contents of sand samples were constantly low. Fe$_o$ contents differed extremely for basin soils, especially for the first 2 depth intervals. Both basin and fringe soils showed an increase of metal oxides within the first depth intervals and in turn a decrease within the deeper layers.

\section{Discussion}
Our results confirmed previously observed differences in \%OC, bulk density and \%TOC between the basin and the fringe mangrove soils \citep{Bolivar2015}.  Furthermore, our measurements of carbon isotopes and mineralogy helped us to establish the potential origin of the carbon stored in both mangrove types. In the following we will discuss these difference and the potential implications of our findings.

>>>>

Different techniques have been used in the past to determine different sources of organic carbon in water saturated ecosystems. For instance, the $\delta^{13}$C isotope has been used to explore different sources of carbon by source partitioning analyses \citep{Bouillion, Spohn2012, Spohn2013}. Also, measurements of the mineralogical composition of the sediments can also give some clues on whether the minerals present in an area are from terrestrial or marine origin.

Here, our main objective was to determine the origin of the relatively high carbon levels in these soils using elemental and isotopic analyses of carbon as well as analyses of the soils' mineralogy in concert with Al- and Fe-oxide measurements. In particular, we expect that: 1) the isotopic composition ($\updelta ^{13}$C) of soil carbon provides information on whether riverine sediments are a main source of C in Cisipat\'a bay, Colombia; and  2) the mineralogical composition of the sediments provides additional information on the origin of the soil carbon and the potential that it is mostly stabilized on the surface of Fe- and Al-oxides. Our main hypothesis is that soils in the basin mangroves are composed mainly by sediments transported by the Sin\'u river and deposited in the delta region, therefore explaining the relatively large values of TOC stored in these soils.

@@

\subsection{Field sampling}
Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}. We sampled 10 plots of 500 m$^2$ up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter (Eijkelkamp bi-partite gouge auger 04.03, Giesbeek, The Netherlands). We selected five randomly chosen points within each plot to extract soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4). In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig. \ref{fig:1}). These river and beach sediment samples were used as end-members in our isotopic analysis. Generally, cores without layer changes were divided into sections every 20 cm to explore differences in carbon concentration with depth. We divided conspicuous layer changes at the boundary, except for the sand and river sediment samples that were analyzed as one single sample for mineralogical and elemental composition.

@@

%\paragraph{X-ray diffraction analysis}
We conducted X-ray diffraction (XRD) measurements for qualitative and quantitative phase analyses \citep{spiess2009} on 12 representative samples including 4 end-member samples, and 4 samples of basin and fringe mangroves each. Samples were measured 20 min each, from 5 to 60 $^{\circ}2\uptheta$. We determined each mineral phase using the powder diffraction file (PDF) data. We also conducted a Rietveld refinement analysis using the Topas software (Bruker Corporation). To define the type of clays included in the samples, we further measured each sample before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels. XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.

@@

\subsection{Data analysis}
We performed statistical tests to determine differences in organic carbon concentrations, total organic carbon, $\delta^{13}$C, and Fe- and Al-oxides, among the fringe and basin mangroves. After an initial test for normality (Shapiro-Wilk test), we found little evidence to support the normality assumption required in common tools such as ANOVA or t-tests. We therefore, performed the non-parametric Mann-Whitney U test, which is equivalent to the two-sample Wilcoxon test \citep{Hollander2015} under the null hypothesis that the samples from the fringe and basin mangroves  differ by a shift location $\mu =0$, and the alternative hypothesis that they are different by some other shift location. In other words, the test helps to determine whether the samples from the two different mangrove types belong to the same statistical distribution as stated in the null hypothesis. Data and code to reproduce all results presented here can be obtained from the following repository \url{https://github.com/crlsierra/mangroveCstabilization.git}.

@@

The different bulk densities between the two forest types, multiplied by their respective \% organic carbon, resulted in a completely different distribution of TOC compared to the previous results based on \%OC alone (Table \ref{tab:1}). A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}). Summed across the profile, fringe mangroves had a TOC of 387.77 $\pm$ 125.77 Mg C ha$^{-1}$, and basing mangroves an average TOC of 665.41 $\pm$ 434.65 Mg C ha$^{-1}$. Given this large variability, and deespite average TOC being almost twice as high in the basing than in the fringe mangroves, we did not find enough evidence to reject the hypothesis that TOC from these two forest types are sampled from the same distribution  ($p$-value $= 0.47$, $W = 350$). TOC decreased with soil depth within the first four (basin) and three (fringe) layers. There was an increase in TOC for the last measured depth intervals of basin and fringe soils, which can be traced back to higher \%OC values of those layers.

We measured a higher proportion of negative $\updelta^{13}$C values for fringe soils \mbox{(-28 to -29 \permil)} than for basin mangroves soils \mbox{(-25 to -29 \permil)} (Fig. \ref{fig:3}).  Depth intervals 20-40 and 40-60 cm of basin soils show outliers with equal $^{13}$C values compared to the Sin\'{u} river end-member samples with values of around \mbox{-25 \permil}. The sand end-member samples have the most positive $^{13}$C values with approximately -3 \permil.  Strong variations in $^{13}$C are conspicuous for the depth intervals 20-40 and 40-60 cm of basin mangrove boxplots. We found no identifiable continuous trend between $^{13}$C values and depth. However, differences in $^{13}$C mean values between fringe and basin mangroves suggest two different distributions ($p$-value $< 0.001$, $W = 542$).

There were no differences between basin and fringe mangroves in terms of the composition of their mineral fraction (Table \ref{tab:2}). The proportion of mineral soil is larger in basin than in fringe mangroves. Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.

@@

The oxalate extraction dissolved much of the poorly crystalline Fe and Al oxides from the amorphous materials, whereas the dithionite extraction dissolved the crystalline Fe oxides as well as the amorphous materials \mbox{(Fig. \ref{fig:6})}. Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $= 0.002$, $W = 325$). Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $W = 364$).
Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values. While riverine samples had similar Al and Fe concentrations than those of basin mangroves in all 4 extraction patterns, metal contents of sand samples were constantly low. Fe$_o$ contents differed extremely for basin soils, especially for the first 2 depth intervals. Both basin and fringe soils showed an increase of metal oxides within the first depth intervals and in turn a decrease within the deeper layers.

\section{Discussion}
Our results confirmed previously observed differences in \%OC, bulk density and \%TOC between the basin and the fringe mangrove soils \citep{Bolivar2015}.  Furthermore, our measurements of carbon isotopes and mineralogy helped us to explore the potential origin of the carbon stored in both mangrove types. In the following, we will discuss these difference and the potential implications of our findings.

### 2018-06-29 Reviews/Response2.tex
Worked on manuscript text

{\color{blue} Thanks for pointing this our. We originally applied an ANOVA test without checking the normality assumption. After checking for normality using the Shapiro-Wilk test, we found that the normally assumption was not met by our data, even after trying different transformations. Therefore, we decided to apply a non-parametric test, the Mann-Whitney U test to determine whether the two independent samples from the fringe and basin mangroves are from the same distributions. After running this test for all the measured variables, we found the same results as with the initial comparison using ANOVA. Therefore, our conclusions are unchanged despite applying a different statistical test. }

@@

>>>>

{\color{blue} Thanks for pointing this out. We originally applied an ANOVA test without checking the normality assumption. After checking for normality using the Shapiro-Wilk test, we found that the normally assumption was not met by our data, even after trying different transformations. Therefore, we decided to apply a non-parametric test, the Mann-Whitney U test to determine whether the two independent samples from the fringe and basin mangroves are from the same distributions. After running this test for all the measured variables, we found the same results as with the initial comparison using ANOVA. Therefore, our conclusions are unchanged despite applying a different statistical test. }

@@

{\color{blue} Organic C concentrations and total N show a very high correlation among each other (0.96 Pearson correlation coefficient), and the C:N ratio does not differ between the two forest types. Therefore, the information in the total N variable is redundant with respect to the information in organic C. Therefore, we think that adding an analysis of the C:N ratio does not contribute much to the story, as opposed to using other independent variables such as the isotopic composition of carbon, and the mineralogical composition of the sediments. }
